Unrecognized States: A Theory of Self-Determination and Foreign Influence

### 2015-09-07 draft.tex
revise proof of prop 2 for decisive war

Player $p$'s continuation value when the status quo is played for $m-1$ periods before ceding is $0 + \frac{\de^m}{1-\de} \alpha$, whereas the continuation value from war is $\frac{\lambda p_{1s} + \alpha p_{2s}}{1-\de} + \frac{\de}{1-\de}\alpha (1-p_{1s}-p_{2s})$. For player $c$, it is $0 + \frac{\de^m}{1-\de} \beta$ versus $\frac{\nu p_{1s} + \beta p_{2s}}{1-\de} + \frac{\de}{1-\de}\beta (1-p_{1s}-p_{2s})$. Thus the condition for $p$'s willingness to instigate fighting to outweigh $c$'s willingness to deter fighting is
\[
  \textstyle \frac{\lambda p_{1s} + \alpha p_{2s}}{1-\de} - \frac{\de \alpha}{1-\de} (p_{1s}+p_{2s}) + \alpha \sum_{i=1}^{m-1} \de^i > \frac{\nu p_{1s} + \beta p_{2s}}{1-\de} - \frac{\de \beta}{1-\de} (p_{1s}+p_{2s}) + \beta \sum_{i=1}^{m-1} \de^i
\]

If this condition holds, then there will be no investments following sanctions. But it is still not guaranteed that the secessionists will choose to cede. Status Quo is no longer a long-run equilibrium strategy, but even without outside investments, it is possible that Fight could be a more attractive action than Cede. Degrading the status quo payoffs is not sufficient to make the continuation value from Cede the highest available.

To see why, note that the sanctions only have a partial effect in reducing the Fight continuation value through the war lottery, while their entire impact is felt on the Status Quo continuation value. Thus the gap between the Status Quo and Fight continuation values narrows at the same time as the corresponding gap between the Status Quo and Cede continuation values. Although the former gap narrows more slowly, it is likely to be smaller at the outset so sanctions do not necessarily make Cede the most preferred action for player $s$.

Therefore, in order to avoid conflict, sanctions must cause player $s$'s continuation value from playing Status Quo to fall below its continuation value from playing Cede before it falls below the continuation value from playing Fight. \hfill $\blacksquare$

>>>>

Player $p$'s continuation value when the status quo is played for $m-1$ periods before ceding is $0 + \frac{\de^m}{1-\de} \alpha$, whereas the continuation value from war is $\frac{\lambda p_{1s} + \alpha (1-p_{1s})}{1-\de}$. For player $c$, it is $0 + \frac{\de^m}{1-\de} \beta$ versus $\frac{\nu p_{1s} + \beta p_{2s}}{1-\de}$. Thus the condition for $p$'s willingness to instigate fighting to outweigh $c$'s willingness to deter fighting is
\[
  \textstyle \frac{\lambda p_{1s} + \alpha (1-p_{1s})}{1-\de} - \frac{\de^m}{1-\de} \alpha > \frac{\nu p_{1s} + \beta (1-p_{1s})}{1-\de} - \beta \frac{\de^m}{1-\de} \alpha
\]

If this condition holds, then there will be no investments following sanctions. But it is still not guaranteed that the secessionists will choose to cede. Status Quo is no longer a long-run equilibrium strategy, but even without outside investments, it is possible that Fight could be a more attractive action than Cede.

The status quo payoffs will degrade period by period in the absence of investments. Whether the secessionists choose Cede or Fight depends on which strategy yields the higher continuation value. In order to avoid conflict, sanctions must cause player $s$'s continuation value from playing Status Quo to fall below its continuation value from playing Cede before it falls below the continuation value from playing Fight. \hfill $\blacksquare$
